---
title: "Global $CO_{2}$ Emissions in 2023 and beyond"
short: "What Keeling missed all these years"
journal: "AER" # AER, AEJ, PP, JEL
month: "`r format(Sys.Date(), '%m')`"
year: "`r format(Sys.Date(), '%Y')`"
author: ""
vol: 0
issue: 0
header-includes: 
  - '\author{}'
  - '\usepackage{graphicx}'
  - '\usepackage{booktabs}'
  - '\usepackage{wrapfig}'
  - '\usepackage{subcaption}'
output: rticles::aea_article
---

```{r setup, echo=FALSE}
## default to not show code, unless we ask for it.
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, dpi = 1000)
options(digits = 8)
```


```{r load packages, echo = FALSE, message = FALSE}
library(tidyverse)
library(tsibble)
library(latex2exp)
library(dplyr)
library(magrittr)
library(patchwork)
library(lubridate)
library(feasts)
library(forecast)
library(sandwich)
library(stargazer)
if (!"fable" %in% rownames(installed.packages())) {
  install.packages("fable")
}
library(fable)
if (!"car" %in% rownames(installed.packages())) {
  install.packages("car")
}
library(car)
theme_set(theme_minimal())
```
\renewcommand{\columnsep}{10pt}
# Introduction
In follow-up to our 1997 report we continue investigating the trend of rising $CO_2$ levels and whether or not it is likely caused by a larger trend or stochastic effects. 

# Mona Loa $CO_{2}$ Data 
NASA \href{https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_daily_mlo.csv}{provides} daily measurement of [$CO_2$] that we download for our analysis.\hfill\break\indent In April of 2019 the Mauna Loa laboratory updated their equipment to measure $CO_2$ with a new technique called Cavity Ring-Down Spectroscopy (CRDS) in contrast to the prior infrared absorption technique. As such, all data from April 2019 onwards will contain measurements using the new method. Additionally, due to eruptions at the Mauna Loa site in 2022, data from December 2022 onwards are from a site at the Maunakea laboratory.

```{r data download, results='hide'}
url <- "https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_daily_mlo.csv" # URL
file_path <- "./data/raw_daily_data.csv" # Destination file
download.file(url, file_path, mode = "w")
```

```{r convert csv to tsibble}
# read csv into a dataframe
df <- read.table(file_path,
  sep = ",",
  dec = ".",
  numerals = c("no.loss"),
  skip = 48
)
# Rename columns for convenience
names(df) <- c("year", "month", "day", "date_dec", "ppm")
# Create date column
df$date <- make_datetime(
  year = df$year,
  month = df$month,
  day = df$day
)
# Drop unneeded columns
df <- df[5:6]

# Convert to tsibble
df <- df %>%
  mutate(date = as.Date(date)) %>%
  as_tsibble(index = "date")

# Aggregate by weeks:
co2_present <- df %>%
  index_by(week = ~ yearweek(.)) %>%
  summarise(ppm = mean(ppm))

# Fill missing values with the average of the nearest existing ones
co2_present <- co2_present %>%
  fill_gaps(.full = TRUE) %>%
  fill(ppm, .direction = "downup") %>%
  fill(ppm, .direction = "updown")

# Create df we will need for plotting
co2_present_df <- data_frame(co2_present)
co2_present_df <- co2_present_df %>%
  mutate(my_date = as.Date(co2_present_df$week))
```


## Recent Trends in Atmospheric Carbon 
\textcolor{red}{Conduct the same EDA on this data. Describe how the Keeling Curve evolved from 1997 to the present, noting where the series seems to be following similar trends to the series that you "evaluated in 1997" and where the series seems to be following different trends. This EDA can use the same, or very similar tools and views as you provided in your 1997 report.}

Atmospheric carbon is plotted in \autoref{fig:OldModels}, and shows relentless increase in atmospheric $CO_2$ concentration even at the remote Hawaiian island.

```{r read objects from 1997 report}
co2_df <- readRDS("co2_df.df")
df_frcst <- readRDS("df_frcst.df")
df_frcst_polynom <- readRDS("df_frcst_polynom.df")
```

```{r plot the keeling curve}
ts_and_forecast <- ggplot() + 
  geom_line(aes(x = my_date, y = ppm, color = "Observed data"), data = co2_present_df) +
  geom_line(aes(x = my_date, y = mean, color = "SARIMA(1.2.1)(0.1.1)"), data = df_frcst) + 
  geom_line(aes(x = my_date, y = pred_quad, color = "Quadratic model w month"), data = df_frcst_polynom) +
  geom_line(aes(x = my_date, y = pred_lin, color = "Linear model"), data = df_frcst_polynom) +
  labs(x = element_blank(), y = TeX(r'(\[$CO_2$\], ppm)')) + 
  xlim (as.Date("1980-01-01"), as.Date("2025-01-01")) + ylim(325, 425) +

  scale_color_manual(name = "", values = c("Observed data" = "black", 
                                         "Quadratic model w month" = "darkgreen",
                                         "SARIMA(1.2.1)(0.1.1)" = "red",
                                          "Linear model" = "blue")) +
  # Legend text and position
  theme(legend.position=c(.25,.85),
        legend.text=element_text(size=20)) +
  # Axis text
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20),
        axis.title.y=element_text(size=20)) +
  # Lines
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())

ggsave(filename = "./figures/plot2_ts.png", plot = ts_and_forecast, device = "png", units = c("in"), height = 5, width = 10)
```

\begin{figure}[ht]
  \includegraphics[width=.8\linewidth]{./figures/plot2_ts.png}
  \caption{Recent trend in [$CO_2$] and models trained on pre-1997 data\label{fig:OldModels}}
\end{figure} 

```{r EDA plots with de-trending}
my_lag <- 52
co2_present <- mutate(co2_present, deseasoned = difference(ppm, lag = my_lag))
co2_present <- mutate(co2_present, detrended = difference(deseasoned, lag = 1))
co2_diff <- slice(co2_present, my_lag + 2:nrow(co2_present))
co2_diff  <- co2_diff  %>% mutate(index = 1:nrow(co2_diff))
co2_diff  <- co2_diff  %>%
  mutate(my_date = make_datetime(year = year(week), month = month(week), day = day(week)))

diff_ts <- co2_diff %>%
  filter(week > yearweek("1998-01-01")) %>%
  ggplot() +
  aes(x = week, y = detrended) +
  geom_line() +
  labs(x = "", y = "Detrended TS")

acf_plot <- co2_diff %>%
  filter(week > yearweek("1998-01-01")) %>%
  ACF(detrended, type = "correlation") %>%
  autoplot() + labs(x = "lag", y = "ACF")

pacf_plot <- co2_diff %>%
  filter(week > yearweek("1998-01-01")) %>%
  ACF(detrended, type = "partial") %>%
  autoplot() + labs(x = "lag", y = "PACF")

ggsave(filename = "./figures/plot_2_2.pdf", plot = diff_ts, device = "pdf", units = c("in"), height = 4, width = 10)

ggsave(filename = "./figures/plot_2_3.pdf", plot = acf_plot, device = "pdf", units = c("in"), height = 4, width = 5)

ggsave(filename = "./figures/plot_2_4.pdf", plot = pacf_plot, device = "pdf", units = c("in"), height = 4, width = 5)
```

# Models and Forecasts 
In this section we will evaluate performance of historic models and generate new models based on the most recent data. 

## 1997 Models vs realized data 
\autoref{fig:OldModels} demonstrates the recent trends in [$CO_{2}$] along with the models fitted on the data prior to 1997. It is clear that ever accelerating trend continues, with no noticeable slow down or even reduced acceleration. Clearly, accuracy of predictions from the linear model does not justify even electricity spent on fitting this model. The trend is accelerating and linear model grossly underestimates the real concentration. Remarkably, simple quadratic model augmented with months as categories is extremely accurate, capturing both accelerating trend and seasonal variability 25 years into the future! Our best ARIMA model underestimates real observations only slightly and its performance is still unexpetedly good for predictions that far out. Worth noticing that both seasonal models perform extremely well up until approximately 2016, when the real trend seemed to have had an additional boost. At then time ARIMA model started falling behind at an increased rate. That might be a statistical fluke, or might indicate the change in the $CO_{2}$ emissions. 

## Performance of 1997 linear and ARIMA models 
```{r read predictions from previous report}
mydat <- readRDS("predictions.num")
first_time.420 <- mydat[1]
expected.420 <- mydat[2]
last_time.420 <- mydat[3]
first_time.500 <- mydat[4]
expected.500 <- mydat[5]
last_time.500 <- mydat[6]
prediction.2100.lower <- mydat[7]
prediction.2100.expected <- mydat[81]
prediction.2100.upper <- mydat[9]
```
It appears that the first time [$CO_{2}$] exceeded 420ppm was in Feb 2023 and the last time it was that low was in Mar 2023. It is unlikely it will return to this value any time soon. That correlates surprisingly well with out 25 year-old prediction that it will reach this threshold on `r expected.420`. As predicted, we are still far away from crossing 500 ppm threshold. 

```{r}
co2_monthly_avg <- co2_present_df %>%
  mutate(month_year = yearmonth(my_date)) %>%
  group_by(month_year) %>%
  summarise(avg_co2 = mean(ppm, na.rm = TRUE))

co2_monthly_avg <-  co2_monthly_avg %>%
  mutate(my_date = make_datetime(year = year(month_year), month = month(month_year), day = 01))

index_of_1998_pred <- match("1998-01-01", df_frcst$my_date)
index_of_today_pred <- match("2023-01-01", df_frcst$my_date)
index_of_1998_real <- match("1998-01-01", co2_monthly_avg$my_date)
index_of_today_real <- match("2023-01-01", co2_monthly_avg$my_date)

#Comparison of forecasted and actual values
comparison <- data.frame(
  date = df_frcst[index_of_1998_pred:index_of_today_pred,1],
  actual = co2_monthly_avg[index_of_1998_real:index_of_today_real,2],
  predicted = df_frcst[index_of_1998_pred:index_of_today_pred,2]
)
names(comparison) <- c("date", "actual", "predicted")
```


```{r}
#Error metrics
comparison <- comparison %>%
  mutate(sqerr = (actual - predicted)^2)

# Calculate overall RMSE
RMSE <- sqrt(mean(comparison$sqerr, na.rm = TRUE))

# 3. Formal test: Mincer-Zarnowitz regression
# First, shift both predicted and actual values by 360 
# This is to avoid calculating intercept too far from the actual data
comparison <- comparison %>% mutate(pred_md = predicted-365, act_md = actual-365)
MZ_reg <- lm(act_md ~ pred_md, data = comparison)

n_d <- data.frame(pred_md = comparison$pred_md)
comparison <- comparison %>% mutate(modeled = predict(MZ_reg, newdata =n_d),
                                  resid = MZ_reg$residuals)

# Run t-test
df <- df.residual(MZ_reg) # Calculate the degrees of freedom
intercept <- coef(MZ_reg)[1] # Get value for intercept
slope <- coef(MZ_reg)[2]-1 # Get value for slope, adjusted by 1

se_intercept <- sqrt(vcov(MZ_reg)[1, 1])  # Standard error of the intercept
t_stat_inter <- intercept / se_intercept # t-statistic of the intercept

se_slope <- sqrt(vcov(MZ_reg)[2, 2])  # Standard error of the intercept
t_stat_slope <- slope / se_slope # t-statistic of the intercept

# p-value for intercept
p_value_intercept <- 2 * pt(abs(t_stat_inter), df = df, lower.tail = FALSE)

# p-value for slope
p_value_slope <- 2 * pt(abs(t_stat_slope), df = df, lower.tail = FALSE)
```
\begin{wrapfigure}[14]{r}{0.5\textwidth}
    \centering
    \vspace{-\normalbaselineskip}
    \includegraphics[height=11\normalbaselineskip]{./figures/plot2_MZreg.png}
     \caption{Mincer-Zarnowitz regression for 1997 ARIMA model\label{fig:MZreg}}
\end{wrapfigure} 
In this case, an RMSE of `r round(RMSE, digits = 1)` means that our model, on average, makes an error of about `r round(RMSE, digits = 1)`ppm in its predictions. \hfill\break\indent We use Mincer-Zarnowitz regression to quantify predictive power of our model. To avoid extrapolation too far from the data, we shift both predicted and actual data by 365 ppm. Intercept of this regression is `r round(MZ_reg$coefficients[1], digits=1)` and  slope is `r round(MZ_reg$coefficients[2], digits=3)`. Given how close these parameters are to 0 and 1, we conclude that our ARIMA model has high predictive power. However, t-test rejects hypothesis that they are equal to 0 and 1 with p-values of `r p_value_intercept`  and `r p_value_slope` respectively. This is likely due to the high number of data points, that makes small deviations highly significant. 

```{r plot MZ regression}
plot_MZreg <- comparison %>% ggplot() + 
  aes(x = pred_md, y = act_md) +
  geom_point(color = "darkgray", size = 0.7) + 
  geom_line(aes(x = pred_md, y = modeled ))
ggsave(filename = "./figures/plot2_MZreg.png", plot = plot_MZreg, device = "png", units = c("in"), height = 4, width = 4)
```

## Best models on present data
\begin{figure}[ht]
  \includegraphics[width=.8\linewidth]{./figures/plot_2_2.pdf}
  \caption{De-seasoned/-trended series for $CO_2$ concentration\label{fig:notrend_ts}}
\end{figure} 
\begin{figure}[h]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot_2_3.pdf}
    \caption{Recent trend in $CO_2$ concentration\label{fig:notrend_acf}}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot_2_4.pdf}
    \caption{Recent trend in $CO_2$ concentration\label{fig:notrend_pacf}}
  \end{subfigure}
  \caption{Diagnostic plots for de-trended series\label{fig:notrend_plots}}
\end{figure}

In this section we will compare perfromance of various models in terms of their predictive power. For that we will save a small portion of our data, the most recent 2 years, as a test set. We will use the rest of the data (train set) to generate the models.
```{r Split into test and train, echo = TRUE}
split.point <- "2021-06-01"

# Create a training set
co2_train <- co2_diff %>% filter((week > yearweek("1998-01-01")) &
  (week < yearweek("2021-06-01")))

# Cretae a test set
co2_test <- co2_diff %>% filter(week >= yearweek(split.point))
```

```{r}
# Helper function to calculate BIC for a fitted model
find_bic <- function(arima_model, ts ){
log_likelihood <- logLik(arima_model) #Get log-likelihood
p <- length(arima_model$coef) # Get the number of parameters
n <- length(ts) # Get the number of observations
bic <- -2 * log_likelihood + p * log(n) # Calculate BIC
print(bic) # Print the BIC value
}
```

```{r Explore model space, results='hide'}
model.nsa <- co2_train %>%
  model(ARIMA(deseasoned ~ 1 + pdq(0:10, 0:2, 0:10) + PDQ(0, 0, 0),
    ic = "bic",
    stepwise = F, greedy = F
  ))
```

```{r exploratory, results='hide' }
arima_model <- arima(co2_train$ppm,
  order = c(0, 1, 2),
  seasonal = list(order = c(0, 0, 0), period = 52)
)

#find_bic(arima_model,co2_train$ppm)
#arima_model %>% residuals() %>% acf(plot = FALSE) %>% autoplot() + coord_cartesian(xlim = c(1, 20))
```

```{r Make final model, results='hide'}
final_arima <- arima(co2_train$ppm,
  order = c(0, 1, 2),
  seasonal = list(order = c(0, 1, 1), period = 52)
)

dual_arima <- arima(co2_train$ppm,
  order = c(0, 2, 2),
  seasonal = list(order = c(0, 1, 1), period = 52)
)

arima_bic <- find_bic(final_arima, co2_train$ppm)
dual_arima_bic <- find_bic(dual_arima, co2_train$ppm)
```

We first focus on seasonally adjusted train data. Grid search of the model space, similar to what we performed in 1997, yields ARIMA(0.1.2)  as the most suitable model with the lowest BIC. Using this model as a baseline, we shift to non-seasonally adjusted data and compensate by adding a seasonal component to our models. The search restricted to PDQ parameter variations shows that  SARIMA(1.1.2)(0.1.1)[52] has the lowest BIC of `r round(arima_bic, digits =1)`. \autoref{fig:resid_acf} shows that ACF plot for this model residuals hardly has significant values, a good indication of the in-sample performance. Building up on our experience from 1997, we also fitted a double-differencing model SARIMA(1.2.2)(0.1.1) that we expect to have better long-term performance. This model has increased BIC of `r round(dual_arima_bic, digits =1)` and demonstrates virtually the same pattern in residual ACF plot (\autoref{fig:resid_acf_dual}). We believe that short-term increased accuracy might not be relevant and simpler model would be better because of reduced prediction uncertainty. However, for longer-term predictions the additional non-linearity of double-differencing will be beneficial.  

```{r Test residuals, results='hide'}
# Visualize ACF plot of its residuals
resid_acf_plot <- final_arima %>%
  residuals() %>%
  acf(plot = FALSE) %>%
  autoplot() + coord_cartesian(xlim = c(1, 20))
ggsave(filename = "./figures/plot2_arimaacf.png", plot = resid_acf_plot, device = "png", units = c("in"), height = 3, width = 4)

# Visualize ACF plot of its residuals
resid_dual_acf_plot <- dual_arima%>%
  residuals() %>%
  acf(plot = FALSE) %>%
  autoplot() + coord_cartesian(xlim = c(1, 20))
ggsave(filename = "./figures/plot2_dualarimaacf.png", plot = resid_dual_acf_plot, device = "png", units = c("in"), height = 3, width = 4)
```
\begin{figure}[h]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot2_arimaacf.png}
    \caption{SARIMA(1.1.2)(0.1.1)[52] \label{fig:resid_acf}}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot2_dualarimaacf.png}
    \caption{SARIMA(1.2.2)(0.1.1)[52] \label{fig:resid_acf_dual}}
  \end{subfigure}
  \caption{ACF plots for SARIMA model residuals}
\end{figure}

```{r Explore model space on raw data, results='hide'}
model.raw <- co2_train %>%
  model(ARIMA(ppm ~ 1 + pdq(0:10, 0:2, 0:10) + PDQ(0, 0, 0),
    ic = "bic",
    stepwise = F, greedy = F
  ))

arima_raw <- arima(co2_train$ppm,
  order = c(2, 1, 3),
  seasonal = list(order = c(0, 0, 0), period = 0))

arima_raw_bic <- find_bic(arima_raw, co2_train$ppm)

# Visualize ACF plot of its residuals
resid_raw_acf_plot <- arima_raw%>%
  residuals() %>%
  acf(plot = FALSE) %>%
  autoplot() + coord_cartesian(xlim = c(1, 20))
ggsave(filename = "./figures/plot2_rawacf.png", plot = resid_raw_acf_plot, device = "png", units = c("in"), height = 4, width = 4)
```
\begin{wrapfigure}[14]{r}{0.5\textwidth}
    \centering
    \vspace{-\normalbaselineskip}
    \includegraphics[height=11\normalbaselineskip]{./figures/plot2_rawacf.png}
    \caption{ACF plot for residuals of ARIMA(2.1.3) model\label{fig:raw_acf}}
\end{wrapfigure} 

Fitting ARIMA models to raw train data with no seasonal adjustments yields inferior models. Grid search of the model space, restricted to pdq parameters only, yields ARIMA(2.1.3)  as the most suitable model with the lowest BIC. However, its BIC `r round(arima_raw_bic, digits = 1)` is significantly higher than that of SARIMA models (`r round(arima_bic, digits =1)` for SARIMA (1.1.2)(0.1.1)) and ACF plot of its residuals (\autoref{fig:raw_acf}) has more significant peaks. It is not clear what advantage ARIMA would have over SARIMA in this context.

```{r Forecast, results='hide'}
frcst_length <- 52 * (2122 - 2021)
forecast_112011 <- forecast(final_arima, h = frcst_length, level = c(95))
forecast_122011 <- forecast(dual_arima, h = frcst_length, level = c(95))
forecast_213000 <- forecast(arima_raw, h = frcst_length, level = c(95))
```

```{r Forecast plotting}

my_date <- seq(ISOdate(2021, 06, 01), by = "week", length.out = frcst_length)

df_toPlot <- cbind.data.frame(my_date = my_date, 
             mean = forecast_112011$mean, lower = forecast_112011$lower[,1], upper = forecast_112011$upper[,1], 
             mean_d = forecast_122011$mean, lower_d = forecast_122011$lower[,1], upper_d = forecast_122011$upper[,1],
             mean_s = forecast_213000$mean, lower_s = forecast_213000$lower[,1], upper_s = forecast_213000$upper[,1])
frcst_arima <- ggplot() +
  geom_line(  aes(x = my_date, y = ppm, color = "Observed data"), data = co2_test)+
  geom_line(  aes(x = my_date, y = mean, color = "SARIMA(1.1.2)(0.1.1)"), data = df_toPlot) +
  geom_ribbon(aes(x = my_date, ymin = lower, ymax = upper), fill = "yellow", alpha = 0.25, data = df_toPlot) +
  geom_line(  aes(x = my_date, y = mean_d, color = "SARIMA(1.2.2)(0.1.1)"), data = df_toPlot) +
  geom_line(  aes(x = my_date, y = mean_s, color = "ARIMA(2.1.3)"), data = df_toPlot) +
  labs(x = element_blank(), y = TeX(r'(\[$CO_2$\], ppm)')) + ylim(410, 430) + 
  xlim (ISOdate(2021, 06, 01), ISOdate(2023, 06, 01)) +
   # Legend colors
  scale_color_manual(name = "", values = c("Observed data" = "black", 
                                         "SARIMA(1.1.2)(0.1.1)" = "orange",
                                         "SARIMA(1.2.2)(0.1.1)" = "red",
                                         "ARIMA(2.1.3)" = "darkgreen")) +
  # Legend text and position
  theme(legend.position=c(.20,.85),
        legend.text=element_text(size=20)) +
  # Axis text
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20),
        axis.title.y=element_text(size=20)) +
  # Lines
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())

ggsave(filename = "./figures/plot2_frcst_arima.png", plot = frcst_arima, device = "png", units = c("in"), height = 4, width = 10)
```

\begin{figure}[ht]
  \includegraphics[width=.8\linewidth]{./figures/plot2_frcst_arima.png}
  \caption{Forecast for future CO2 concentration \label{fig:frcst2y}}
\end{figure}
Predictive performance of the three models discussed above is illustrated by \autolabel{fig:frcst2y}.




```{r qaud on deseasoned data}
model.quad <- lm(formula = deseasoned ~ index + I(index^2), data = co2_train)
arima_deseasoned <- arima(co2_train$deseasoned,
  order = c(0, 1, 2),
  seasonal = list(order = c(0, 0, 0), period = 52))
```
Fitting quadratic model to de-seasoned data reveals trends similar to the ones we saw in 1997. 

## Forecasts: How bad could it get?

\textcolor{red}{With the non-seasonally adjusted data series, generate predictions for when atmospheric CO2 is expected to be at 420 ppm and 500 ppm levels for the first and final times (consider prediction intervals as well as point estimates in your answer). Generate a prediction for atmospheric CO2 levels in the year 2122. How confident are you that these will be accurate predictions?}


# Conclusions 

\textcolor{red}{What to conclude is unclear. }

\bibliographystyle{aea}
\bibliography{references}

\appendix
\section{Appendix: Model Robustness}

\textcolor{red}{While the most plausible model that we estimate is reported in the main, "Modeling" section, in this appendix to the article we examine alternative models. Here, our intent is to provide a skeptic that does not accept our assessment of this model as an ARIMA of order (1,2,3) an understanding of model forecasts under alternative scenarios. }
