---
title: "Global $CO_{2}$ Emissions in 2023 and beyond"
short: "What Keeling missed all these years"
journal: "AER" # AER, AEJ, PP, JEL
month: "`r format(Sys.Date(), '%m')`"
year: "`r format(Sys.Date(), '%Y')`"
vol: 0
issue: 0
author:
  - name: Sean Seneviratne
    firstname: Sean
    surname: Seneviratne
    email: sean_senev@ischool.berkeley.edu
    affiliation: UC Berkeley
  - name: Trevor Dalton
    firstname: Trevor
    surname: Dalton
    email: trevor.dalton@ischool.berkeley.edu
    affiliation: UC Berkeley
  - name: Kyle Ruan
    firstname: Kyle
    surname: Ruan
    email: kyleruan@ischool.berkeley.edu
    affiliation: UC Berkeley
  - name: Artem Lebedev
    firstname: Artem 
    surname: Lebedev
    email: artem.lebedev@berkeley.edu
    affiliation: McMaster University
header-includes: 
  - '\author{}'
  - '\usepackage{graphicx}'
  - '\usepackage{booktabs}'
  - '\usepackage{wrapfig}'
  - '\usepackage{subcaption}'
output: rticles::aea_article
---

```{r setup, echo=FALSE}
## default to not show code, unless we ask for it.
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, dpi = 600)
options(digits = 8)
```


```{r load packages, echo = FALSE, message = FALSE}
library(tidyverse)
library(tsibble)
library(latex2exp)
library(dplyr)
library(magrittr)
library(patchwork)
library(lubridate)
library(feasts)
library(forecast)
library(sandwich)
library(stargazer)
if (!"fable" %in% rownames(installed.packages())) {
  install.packages("fable")
}
library(fable)
if (!"car" %in% rownames(installed.packages())) {
  install.packages("car")
}
library(car)
theme_set(theme_minimal())
```
\renewcommand{\columnsep}{10pt}
# Introduction
\begin{figure}[ht]
  \includegraphics[width=.8\linewidth]{./figures/plot2_ts.png}
  \caption{Recent trend in [$CO_2$] and models trained on pre-1997 data\label{fig:OldModels}}
\end{figure} 
In follow-up to our 1997 report we continue investigating the trend of rising $CO_2$ levels and whether or not it is likely caused by a larger trend or stochastic effects. 

# Mona Loa $CO_{2}$ Data 
NASA \href{https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_daily_mlo.csv}{provides} daily measurement of [$CO_2$] that we download for our analysis.\hfill\break\indent In April of 2019 the Mauna Loa laboratory updated their equipment to measure $CO_2$ with a new technique called Cavity Ring-Down Spectroscopy (CRDS) in contrast to the prior infrared absorption technique. As such, all data from April 2019 onwards will contain measurements using the new method. Additionally, due to eruptions at the Mauna Loa site in 2022, data from December 2022 onwards are from a site at the Maunakea laboratory.

```{r data download, results='hide'}
url <- "https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_daily_mlo.csv" # URL
file_path <- "./data/raw_daily_data.csv" # Destination file
download.file(url, file_path, mode = "w")
```

```{r convert csv to tsibble}
# read csv into a dataframe
df <- read.table(file_path,
  sep = ",",
  dec = ".",
  numerals = c("no.loss"),
  skip = 48
)
# Rename columns for convenience
names(df) <- c("year", "month", "day", "date_dec", "ppm")
# Create date column
df$date <- make_datetime(
  year = df$year,
  month = df$month,
  day = df$day
)
# Drop unneeded columns
df <- df[5:6]

# Convert to tsibble
df <- df %>%
  mutate(date = as.Date(date)) %>%
  as_tsibble(index = "date")

# Aggregate by weeks:
co2_present <- df %>%
  index_by(week = ~ yearweek(.)) %>%
  summarise(ppm = mean(ppm))

# Fill missing values with the average of the nearest existing ones
co2_present <- co2_present %>%
  fill_gaps(.full = TRUE) %>%
  fill(ppm, .direction = "downup") %>%
  fill(ppm, .direction = "updown")

# Create df we will need for plotting
co2_present_df <- data_frame(co2_present)
co2_present_df <- co2_present_df %>%
  mutate(my_date = as.Date(co2_present_df$week))
```


## Recent Trends in Atmospheric Carbon 
\begin{figure}[h]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot2_deseasoned_ts.png}
    \setlength{\belowcaptionskip}{-5pt}
    \caption{De-seasoned series\label{fig:notrend_ts}}
  \end{subfigure}
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot2_acf_ns.png}
    \caption{ACF\label{fig:notrend_acf}}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot2_pacf_ns.png}
    \caption{PACF\label{fig:notrend_pacf}}
  \end{subfigure}
  \setlength{\belowcaptionskip}{-10pt}
  \caption{Diagnostic plots for de-seasoned series of [$CO_2$] after 1996\label{fig:notrend_plots}}
\end{figure}

Atmospheric $CO_{2}$ level is plotted in \autoref{fig:OldModels}(black trace), and shows relentless increase even at the remote Hawaiian island. In fact, the $CO_{2}$ emissions level is rising at a higher rate between 2010 and now than the period before 1996, as demonstrated by the steepening slope in the chart below. In general, the trend is similar to the trend observed before 1996: steady uptrend.\hfill\break\indent After removing yearly variability from weekly data by differencing with lag 52 the series becomes close to stationary, although with some upward trend (\autoref{fig:notrend_ts}). ACF plot (\autoref{fig:notrend_acf}) of the series shows gradual decay and PACF plot (\autoref{fig:notrend_pacf}) has a few significant peaks up to lag 4. These features are indicative of AR process, but presence of upward trend complicates the analysis. A grid search of model parameters, including differencing parameter will be required for precise model allocation. 

```{r read objects from 1997 report}
co2_df <- readRDS("co2_df.df")
df_frcst <- readRDS("df_frcst.df")
df_frcst_polynom <- readRDS("df_frcst_polynom.df")
```

```{r plot the keeling curve}
ts_and_forecast <- ggplot() + 
  geom_line(aes(x = my_date, y = ppm, color = "Observed data"), data = co2_present_df) +
  geom_line(aes(x = my_date, y = mean, color = "SARIMA(1.2.1)(0.1.1)"), data = df_frcst) + 
  geom_line(aes(x = my_date, y = pred_quad, color = "Quadratic model w month"), data = df_frcst_polynom) +
  geom_line(aes(x = my_date, y = pred_lin, color = "Linear model"), data = df_frcst_polynom) +
  labs(x = element_blank(), y = TeX(r'(\[$CO_2$\], ppm)')) + 
  xlim (as.Date("1980-01-01"), as.Date("2025-01-01")) + ylim(325, 425) +

  scale_color_manual(name = "", values = c("Observed data" = "black", 
                                         "Quadratic model w month" = "darkgreen",
                                         "SARIMA(1.2.1)(0.1.1)" = "red",
                                          "Linear model" = "blue")) +
  # Legend text and position
  theme(legend.position=c(.25,.85),
        legend.text=element_text(size=20)) +
  # Axis text
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20),
        axis.title.y=element_text(size=20)) +
  # Lines
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())

ggsave(filename = "./figures/plot2_ts.png", plot = ts_and_forecast, device = "png", units = c("in"), height = 5, width = 10)
```

```{r EDA plots with de-trending}
my_lag <- 52
co2_present <- mutate(co2_present, deseasoned = difference(ppm, lag = my_lag))
co2_present <- mutate(co2_present, detrended = difference(deseasoned, lag = 1))
co2_diff <- slice(co2_present, my_lag + 2:nrow(co2_present))
co2_diff  <- co2_diff  %>% mutate(index = 1:nrow(co2_diff))
co2_diff  <- co2_diff  %>%
  mutate(my_date = make_datetime(year = year(week), month = month(week), day = day(week)))

diff_ts <- co2_diff %>%
  filter(week > yearweek("1996-01-01")) %>%
  ggplot() +
  aes(x = my_date, y = deseasoned) +
  geom_line() +
  labs(x = "", y = "Deseasoned TS")+
  # Axis text
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20),
        axis.title.x=element_text(size=20),
        axis.title.y=element_text(size=20))

acf_plot <- co2_diff %>%
  filter(week > yearweek("1996-01-01")) %>%
  ACF(deseasoned, type = "correlation") %>%
  autoplot() + labs(x = "lag", y = "ACF")+
  # Axis text
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20),
        axis.title.y=element_text(size=20))

pacf_plot <- co2_diff %>%
  filter(week > yearweek("1996-01-01")) %>%
  ACF(deseasoned, type = "partial") %>%
  autoplot() + labs(x = "lag", y = "PACF")+
  # Axis text
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20),
        axis.title.y=element_text(size=20))

ggsave(filename = "./figures/plot2_deseasoned_ts.png", plot = diff_ts, device = "png", units = c("in"), height = 3, width = 8)
ggsave(filename = "./figures/plot2_acf_ns.png", plot = acf_plot, device = "png", units = c("in"), height = 3, width = 4)
ggsave(filename = "./figures/plot2_pacf_ns.png", plot = pacf_plot, device = "png", units = c("in"), height = 3, width = 4)

```

# Models and Forecasts 
In this section we will evaluate performance of historic models and generate new models based on the most recent data. 

## 1997 Models vs realized data 
\autoref{fig:OldModels} demonstrates the recent trends in [$CO_{2}$] along with the models fitted on the data prior to 1997. It is clear that ever accelerating trend continues, with no noticeable slow down or even reduced acceleration. Clearly, accuracy of predictions from the linear model does not justify even electricity spent on fitting this model. The trend is accelerating and linear model grossly underestimates the real concentration. Remarkably, simple quadratic model augmented with months as categories is extremely accurate, capturing both accelerating trend and seasonal variability 25 years into the future! Our best ARIMA model underestimates real observations only slightly and its performance is still unexpetedly good for predictions that far out. Worth noticing that both seasonal models perform extremely well up until approximately 2016, when the real trend seemed to have had an additional boost. At then time ARIMA model started falling behind at an increased rate. That might be a statistical fluke, or might indicate the change in the $CO_{2}$ emissions. 

## Performance of 1997 linear and ARIMA models 
\begin{wrapfigure}[16]{r}{0.5\textwidth}
    \centering
    \includegraphics[height=11\normalbaselineskip]{./figures/plot2_MZreg.png}
    \caption{Mincer-Zarnowitz regression for 1997 ARIMA model\label{fig:MZreg}}
\end{wrapfigure} 

```{r read predictions from previous report}
mydat <- readRDS("predictions.num")
first_time.420 <- mydat[1]
expected.420 <- mydat[2]
last_time.420 <- mydat[3]
first_time.500 <- mydat[4]
expected.500 <- mydat[5]
last_time.500 <- mydat[6]
prediction.2100.lower <- mydat[7]
prediction.2100.expected <- mydat[81]
prediction.2100.upper <- mydat[9]
```
It appears that the first time [$CO_{2}$] exceeded 420 ppm was in Feb 2022 and the last time it was that low was in Mar 2023. It is unlikely it will return to this value any time soon. That correlates surprisingly well with out 25 year-old prediction that it will reach this threshold on `r expected.420`. As predicted, we are still far away from crossing 500 ppm threshold. 

```{r}
co2_monthly_avg <- co2_present_df %>%
  mutate(month_year = yearmonth(my_date)) %>%
  group_by(month_year) %>%
  summarise(avg_co2 = mean(ppm, na.rm = TRUE))

co2_monthly_avg <-  co2_monthly_avg %>%
  mutate(my_date = make_datetime(year = year(month_year), month = month(month_year), day = 01))

index_of_1998_pred <- match("1998-01-01", df_frcst$my_date)
index_of_today_pred <- match("2023-01-01", df_frcst$my_date)
index_of_1998_real <- match("1998-01-01", co2_monthly_avg$my_date)
index_of_today_real <- match("2023-01-01", co2_monthly_avg$my_date)

#Comparison of forecasted and actual values
comparison <- data.frame(
  date = df_frcst[index_of_1998_pred:index_of_today_pred,1],
  actual = co2_monthly_avg[index_of_1998_real:index_of_today_real,2],
  predicted = df_frcst[index_of_1998_pred:index_of_today_pred,2]
)
names(comparison) <- c("date", "actual", "predicted")
```


```{r}
#Error metrics
comparison <- comparison %>%
  mutate(sqerr = (actual - predicted)^2)

# Calculate overall RMSE
RMSE <- sqrt(mean(comparison$sqerr, na.rm = TRUE))

# 3. Formal test: Mincer-Zarnowitz regression
# First, shift both predicted and actual values by 360 
# This is to avoid calculating intercept too far from the actual data
comparison <- comparison %>% mutate(pred_md = predicted-365, act_md = actual-365)
MZ_reg <- lm(act_md ~ pred_md, data = comparison)

n_d <- data.frame(pred_md = comparison$pred_md)
comparison <- comparison %>% mutate(modeled = predict(MZ_reg, newdata =n_d),
                                  resid = MZ_reg$residuals)

# Run t-test
df <- df.residual(MZ_reg) # Calculate the degrees of freedom
intercept <- coef(MZ_reg)[1] # Get value for intercept
slope <- coef(MZ_reg)[2]-1 # Get value for slope, adjusted by 1

se_intercept <- sqrt(vcov(MZ_reg)[1, 1])  # Standard error of the intercept
t_stat_inter <- intercept / se_intercept # t-statistic of the intercept

se_slope <- sqrt(vcov(MZ_reg)[2, 2])  # Standard error of the intercept
t_stat_slope <- slope / se_slope # t-statistic of the intercept

# p-value for intercept
p_value_intercept <- 2 * pt(abs(t_stat_inter), df = df, lower.tail = FALSE)

# p-value for slope
p_value_slope <- 2 * pt(abs(t_stat_slope), df = df, lower.tail = FALSE)
```

In this case, an RMSE of `r round(RMSE, digits = 1)` means that our model, on average, makes an error of about `r round(RMSE, digits = 1)` ppm in its predictions. \hfill\break\indent We use Mincer-Zarnowitz regression to quantify predictive power of our model. To avoid extrapolation too far from the data, we shift both predicted and actual data by 365 ppm. Intercept of this regression is `r round(MZ_reg$coefficients[1], digits=1)` and  slope is `r round(MZ_reg$coefficients[2], digits=3)`. Given how close these parameters are to 0 and 1, we conclude that our ARIMA model has high predictive power. t-test fails to reject hypothesis that intercept is not 0 with p value `r round(p_value_intercept, digits = 2)`. However, t-test rejects hypothesis that slope is 1 with p-value `r p_value_slope`. This is likely due to the high number of data points, that makes small deviations highly significant. 

```{r plot MZ regression}
plot_MZreg <- comparison %>% ggplot() + 
  aes(x = pred_md, y = act_md) +
  geom_point(color = "darkgray", size = 0.7) + 
  geom_line(aes(x = pred_md, y = modeled )) +
  labs(x = "Predicted, shifted", y = "Actual, shifted") + 
  # Axis text
  theme(axis.text.x=element_text(size=12), 
        axis.text.y=element_text(size=12), 
        axis.title.x=element_text(size=12),
        axis.title.y=element_text(size=12))
ggsave(filename = "./figures/plot2_MZreg.png", plot = plot_MZreg, device = "png", units = c("in"), height = 3, width = 4)
```

## Best models on present data
In this section we will compare performance of various models in terms of their predictive power. For that we will save a small portion of our data, after June 2021, as a test set. We will use the data from 1998 to 2021 as a training set to generate the models. We decided to not use pre-1998 data as there were at least two areas of irregularity: around gas price hike early in the 1970 and around Asian economic crisis of the 1990's.
```{r Split into test and train}
split.point <- "2021-06-01"

# Create a training set
co2_train <- co2_diff %>% filter((week > yearweek("1998-01-01")) &
  (week < yearweek("2021-06-01")))

# Cretae a test set
co2_test <- co2_diff %>% filter(week >= yearweek(split.point))
```

```{r}
# Helper function to calculate BIC for a fitted model
find_bic <- function(arima_model, ts ){
log_likelihood <- logLik(arima_model) #Get log-likelihood
p <- length(arima_model$coef) # Get the number of parameters
n <- length(ts) # Get the number of observations
bic <- -2 * log_likelihood + p * log(n) # Calculate BIC
print(bic) # Print the BIC value
}
```

```{r Explore model space, results='hide'}
model.nsa <- co2_train %>%
  model(ARIMA(deseasoned ~ 1 + pdq(0:10, 0:2, 0:10) + PDQ(0, 0, 0),
    ic = "bic",
    stepwise = F, greedy = F
  ))
```

```{r exploratory, results='hide' }
arima_model <- arima(co2_train$ppm,
  order = c(0, 1, 2),
  seasonal = list(order = c(0, 0, 0), period = 52)
)

#find_bic(arima_model,co2_train$ppm)
#arima_model %>% residuals() %>% acf(plot = FALSE) %>% autoplot() + coord_cartesian(xlim = c(1, 20))
```

```{r Make final model, results='hide'}
final_arima <- arima(co2_train$ppm,
  order = c(0, 1, 2),
  seasonal = list(order = c(0, 1, 1), period = 52)
)

dual_arima <- arima(co2_train$ppm,
  order = c(0, 2, 2),
  seasonal = list(order = c(0, 1, 1), period = 52)
)

arima_bic <- find_bic(final_arima, co2_train$ppm)
dual_arima_bic <- find_bic(dual_arima, co2_train$ppm)
```

We first focus on seasonally adjusted train data. Grid search of the model space, similar to what we performed in 1997, yields ARIMA(0.1.2)  as the most suitable model with the lowest BIC. Using this model as a baseline, we shift to non-seasonally adjusted data and compensate by adding a seasonal component to our models. The search restricted to PDQ parameter variations shows that  SARIMA(1.1.2)(0.1.1)[52] has the lowest BIC of `r round(arima_bic, digits =1)`. \autoref{fig:resid_acf} shows that ACF plot for this model residuals hardly has significant values, a good indication of the in-sample performance. Building up on our experience from 1997, we also fitted a double-differencing model SARIMA(1.2.2)(0.1.1) that we expect to have better long-term performance. This model has increased BIC of `r round(dual_arima_bic, digits =1)` and demonstrates virtually the same pattern in residual ACF plot (\autoref{fig:resid_acf_dual}). We believe that short-term increased accuracy might not be relevant and simpler model would be better because of reduced prediction uncertainty. However, for longer-term predictions the additional non-linearity of double-differencing will be beneficial.  

```{r Test residuals, results='hide'}
# Visualize ACF plot of its residuals
resid_acf_plot <- final_arima %>%
  residuals() %>%
  acf(plot = FALSE) %>%
  autoplot() + coord_cartesian(xlim = c(1, 20))
ggsave(filename = "./figures/plot2_arimaacf.png", plot = resid_acf_plot, device = "png", units = c("in"), height = 3, width = 4)

# Visualize ACF plot of its residuals
resid_dual_acf_plot <- dual_arima%>%
  residuals() %>%
  acf(plot = FALSE) %>%
  autoplot() + coord_cartesian(xlim = c(1, 20))
ggsave(filename = "./figures/plot2_dualarimaacf.png", plot = resid_dual_acf_plot, device = "png", units = c("in"), height = 3, width = 4)
```
\begin{figure}[h]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot2_arimaacf.png}
    \caption{SARIMA(1.1.2)(0.1.1)[52] \label{fig:resid_acf}}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot2_dualarimaacf.png}
    \caption{SARIMA(1.2.2)(0.1.1)[52] \label{fig:resid_acf_dual}}
  \end{subfigure}
  \caption{ACF plots for SARIMA model residuals}
\end{figure}

```{r Explore model space on raw data, results='hide'}
model.raw <- co2_train %>%
  model(ARIMA(ppm ~ 1 + pdq(0:10, 0:2, 0:10) + PDQ(0, 0, 0),
    ic = "bic",
    stepwise = F, greedy = F
  ))

arima_raw <- arima(co2_train$ppm,
  order = c(2, 1, 3),
  seasonal = list(order = c(0, 0, 0), period = 0))

arima_raw_bic <- find_bic(arima_raw, co2_train$ppm)

# Visualize ACF plot of its residuals
resid_raw_acf_plot <- arima_raw%>%
  residuals() %>%
  acf(plot = FALSE) %>%
  autoplot() + coord_cartesian(xlim = c(1, 20))
ggsave(filename = "./figures/plot2_rawacf.png", plot = resid_raw_acf_plot, device = "png", units = c("in"), height = 4, width = 4)
```
\begin{wrapfigure}[14]{r}{0.5\textwidth}
    \centering
    \vspace{-\normalbaselineskip}
    \includegraphics[height=11\normalbaselineskip]{./figures/plot2_rawacf.png}
    \caption{ACF plot for residuals of ARIMA(2.1.3) model\label{fig:raw_acf}}
\end{wrapfigure} 

Fitting ARIMA models to raw train data with no seasonal adjustments yields inferior models. Grid search of the model space, restricted to pdq parameters only, yields ARIMA(2.1.3)  as the most suitable model with the lowest BIC. However, its BIC `r round(arima_raw_bic, digits = 1)` is significantly higher than that of SARIMA models (`r round(arima_bic, digits =1)` for SARIMA (1.1.2)(0.1.1)) and ACF plot of its residuals (\autoref{fig:raw_acf}) has more significant peaks. It is not clear what advantage ARIMA would have over SARIMA in this context.

```{r Forecast, results='hide'}
frcst_length <- 52 * (2122 - 2021)
forecast_112011 <- forecast(final_arima, h = frcst_length, level = c(95))
forecast_122011 <- forecast(dual_arima, h = frcst_length, level = c(95))
forecast_213000 <- forecast(arima_raw, h = frcst_length, level = c(95))
```

```{r Forecast plotting}
my_date <- seq(ISOdate(2021, 06, 01), by = "week", length.out = frcst_length)

df_toPlot <- cbind.data.frame(my_date = my_date, 
             mean = forecast_112011$mean, lower = forecast_112011$lower[,1], upper = forecast_112011$upper[,1], 
             mean_d = forecast_122011$mean, mean_s = forecast_213000$mean)
frcst_arima <- ggplot() +
  geom_line(  aes(x = my_date, y = ppm, color = "Observed data"), data = co2_test)+
  geom_line(  aes(x = my_date, y = mean, color = "SARIMA(1.1.2)(0.1.1)"), data = df_toPlot) +
  geom_ribbon(aes(x = my_date, ymin = lower, ymax = upper), fill = "yellow", alpha = 0.25, data = df_toPlot) +
  geom_line(  aes(x = my_date, y = mean_d, color = "SARIMA(1.2.2)(0.1.1)"), data = df_toPlot) +
  geom_line(  aes(x = my_date, y = mean_s, color = "ARIMA(2.1.3)"), data = df_toPlot) +
  labs(x = element_blank(), y = TeX(r'(\[$CO_2$\], ppm)')) + ylim(410, 430) + 
  xlim (ISOdate(2021, 06, 01), ISOdate(2023, 06, 01)) +
   # Legend colors
  scale_color_manual(name = "", values = c("Observed data" = "black", 
                                         "SARIMA(1.1.2)(0.1.1)" = "orange",
                                         "SARIMA(1.2.2)(0.1.1)" = "red",
                                         "ARIMA(2.1.3)" = "darkgreen")) +
  # Legend text and position
  theme(legend.position=c(.30,.85),
        legend.text=element_text(size=20)) +
  # Axis text
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20),
        axis.title.y=element_text(size=20)) +
  # Lines
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())

ggsave(filename = "./figures/plot2_frcst_arima.png", plot = frcst_arima, device = "png", units = c("in"), height = 4, width = 7)
```

\begin{figure}[h]
  \centering
  \begin{subfigure}[b]{0.65\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot2_frcst_arima.png}
    \caption{Raw data \label{fig:frcst2y}}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{./figures/plot2_frcst_nsa.png}
    \caption{Deseasoned data \label{fig:frcst2y_raw}}
  \end{subfigure}
  \setlength{\belowcaptionskip}{-30pt}
  \caption{Predictive performance of the fitted models}
\end{figure}

```{r qaud on deseasoned data}
model.quad <- lm(formula = deseasoned ~ index + I(index^2), data = co2_train)
arima_deseasoned <- arima(co2_train$deseasoned,
  order = c(0, 1, 2),
  seasonal = list(order = c(0, 0, 0), period = 52))

forecast_012 <- forecast(arima_deseasoned, h = frcst_length, level = c(95))
forecast_quad <- fitted(model.quad, h = frcst_length)
idx = length(forecast_quad)

df_toPlot_ns <-  cbind.data.frame(my_date = my_date[1:idx], 
             mean = forecast_012$mean[1:idx], lower = forecast_012$lower[1:idx,1], upper = forecast_012$upper[1:idx,1],
             mean_d = forecast_quad)


```

```{r plotting deseasoned forecast}
frcst_arima_nsa <- ggplot() +
  geom_line(  aes(x = my_date, y = deseasoned, color = "Observed data"), data = co2_test)+
  geom_line(  aes(x = my_date, y = mean, color = "ARIMA(0.1.2)"), data = df_toPlot_ns) +
  geom_ribbon(aes(x = my_date, ymin = lower, ymax = upper), fill = "yellow", alpha = 0.25, data = df_toPlot_ns) +
  geom_line(  aes(x = my_date, y = mean_d, color = "Quadratic"), data = df_toPlot_ns) +
  labs(x = element_blank(), y = TeX(r'(d\[$CO_2$\], ppm)')) + ylim(-1, 6) + 
  xlim (ISOdate(2021, 06, 01), ISOdate(2023, 06, 01)) +
   # Legend colors
  scale_color_manual(name = "", values = c("Observed data" = "black", 
                                         "ARIMA(0.1.2)" = "orange",
                                         "Quadratic" = "darkgreen")) +
  # Legend text and position
  theme(legend.position=c(.40,.85),
        legend.text=element_text(size=20)) +
  # Axis text
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20),
        axis.title.y=element_text(size=20)) +
  # Lines
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())

ggsave(filename = "./figures/plot2_frcst_nsa.png", plot = frcst_arima_nsa, device = "png", units = c("in"), height = 4, width = 3)
```

Predictive performance of the three models discussed above is illustrated by \autoref{fig:frcst2y}. Both SARIMA models closely match observed data. Confidence interval of SARIMA(1.1.2)(0.1.1) shown in yellow, contains test data throughout whole region. ARIMA model, lacking seasonal component, fails almost immediately. Fitting quadratic model to de-seasoned data demonstrates very good performance, as does ARIMA model (\autoref{fig:frcst2y_raw}). 

## Forecasts: How bad could it get?
```{r forecast rising CO2 levels}
# Return the date that the given series first hit x num
first.date <- function(series, num) {
  return(df_toPlot$my_date[which(series > num)[1]])
}

# When CO2 is expected to hit 420ppm
first_time.420 <- as.Date(first.date(df_toPlot$upper, 420))
expected.420 <- as.Date(first.date(df_toPlot$mean, 420))
last_time.420 <- as.Date(first.date(df_toPlot$lower, 420))

# When CO2 is expected to hit 500ppm
first_time.500 <- as.Date(first.date(df_toPlot$upper, 500))
expected.500 <- as.Date(first.date(df_toPlot$mean, 500))
last_time.500 <- as.Date(first.date(df_toPlot$lower, 500))

# Expected CO2 levels in the year 2100
prediction.2122.index <- length(df_toPlot$my_date)
prediction.2122.lower <- df_toPlot$lower[prediction.2122.index]
prediction.2122.expected <- df_toPlot$mean[prediction.2122.index]
prediction.2122.upper <- df_toPlot$upper[prediction.2122.index]
```
\begin{wraptable}[6]{r}{0.6\textwidth}
  \setlength{\belowcaptionskip}{-15pt}
  \caption{[$CO_{2}$] thresholds crossing}\label{wrap-tab:tab1}
  \begin{tabular}{llll}
    \toprule 
    ppm & As early as & Expected & Last time \\
    \hline
    420 & `r first_time.420` & `r expected.420` & `r last_time.420` \\
    500 & `r first_time.500` & `r expected.500` & `r last_time.500`  \\
    \bottomrule
  \end{tabular}
\end{wraptable}

Based on the SARIMA(1.2.2)(0.1.1) model that we deem more suitable for long-term forecasting, we can attempt to predict some important milestones. For the year 2122 we expect $[CO_{2}]$ to be `r round(prediction.2122.expected, 0)` ppm  with 95% CI from `r round(prediction.2122.lower, 0)` ppm to `r round(prediction.2122.upper, 0)` ppm. Our previous model was successful at predicting data 25 years out, and the overall trend remains the same for the past 60 years, so we are cautiously optimistic about quality of this prediction. There is so far no sign of slowing down of the trend, so sigmoidal inflection will not happen any time soon. \autoref{wrap-tab:tab1} summarizes when 420 ppm and 500 ppm thresholds might be crossed. Fast-increasing uncertainty in predictions, typical for SARIMA models, limits our ability to make meaningful predictions about distant future. For instance, despite fast growing trend, the model can not exclude possibility that atmospheric $CO_{2}$ will never exceed 500 ppm thresholds.

# Conclusions 

In this report, we assessed data from the Mona Loa observatory to model and predict atmospheric CO2 concentrations for the next century. Compared to the last report, we have access to present data, which improves our predictive power. The forecast from our modeling is only valid assuming all forces that currently influence atmospheric carbon remain unchanged. Given this
reasonable assumption, our model still predicts a very grim future for the global climate if no actions are taken. In fact, our predictions are fairly close to the last report, which indicates that nothing has been done to stop the trend. The global CO2 level will reach a whopping `r round(prediction.2122.expected, 0)` ppm in 2122 based on 
our forecast, making the planet uninhabitable.
