---
title: 'Statistical Methods for Discrete Response, Time Series, and Panel Data (W271): Lab 2'
geometry: margin=1in
output:
  github_document: default
---

# The Keeling Curve

In the 1950s, the geochemist Charles David Keeling observed a seasonal pattern in the amount of carbon dioxide present in air samples collected over the course of several years. He was able to attribute this pattern to the variation in global rates of photosynthesis throughout the year, caused by the difference in land area and vegetation cover between the Earth's northern and southern hemispheres. 

In 1958 Keeling began continuous monitoring of atmospheric carbon dioxide concentrations from the Mauna Loa Observatory in Hawaii and soon observed a trend increase carbon dioxide levels in addition to the seasonal cycle. He was able to attribute this trend increase to growth in global rates of fossil fuel combustion. This trend has continued to the present, and is known as the "Keeling Curve."

```{r load packages, echo = FALSE, message = FALSE}
install.packages("tidyverse")
library(tsibble)
library(latex2exp) 
library(feasts)
library(forecast)
library(tseries)
library(lubridate)
```

```{r}
library(ggplot2)
```


```{r load packages, echo = FALSE, message = FALSE}
theme_set(theme_minimal())
knitr::opts_chunk$set(dpi=1000)
```

```{r plot the keeling curve, echo = FALSE}
tsibble::as_tsibble(co2) %>%
  ggplot() + 
  aes(x=index, y=value) + 
  geom_line(color = 'steelblue') +
  labs(
    title = TeX(r'(Monthly Mean $CO_2$)'),
    subtitle = 'The "Keeling Curve"',
    x = 'Month and Year',
    y = TeX(r'($CO_2$ parts per million)')
  )
```
\newpage

# Your Assignment 

Your goal in this assignment is to produce a comprehensive analysis of the Mona Loa CO2 data that you will be read by an interested, supervising data scientist. Rather than this being a final report, you might think of this as being a contribution to your laboratory. You and your group have been initially charged with the task of investigating the trends of global CO2, and told that if you find "anything interesting" that the team may invest more resources into assessing the question. 

Because this is the scenario that you are responding to: 

1. Your writing needs to be clear, well-reasoned, and concise. Your peers will be reading this, and you have a reputation to maintain.
2. Decisions that you make for your analysis need also be clear and well-reasoned. While the main narrative of your deliverable might only present the modeling choices that you determine are the most appropriate, there might exist supporting materials that examine what the consequences of other choices would be. As a concrete example, if you determine that a series is an AR(1) process your main analysis might provide the results of the critical test that led you to that determination and the results of the rest of the analysis under AR(1) modeling choices. However, in an appendix or separate document that is linked in your main report, you might show what a MA model would have meant for your results instead.
3. Your code and repository are a part of the deliverable. If you were to make a clear argument that this is a question worth pursuing, but then when the team turned to continue the work they found a repository that was a jumble of coding idioms, version-ed or outdated files, and skeletons it would be a disappointment.

# Report from the Point of View of 1997 

For the first part of this task, suspend reality for a short period of time and conduct your analysis from the point of view of a data scientist doing their work in the early months of 1998. Do this by using data that is included in _every_ R implementation, the `co2` dataset. This dataset is lazily loaded with every R instance, and is stored in an object called `co2`. 

## (3 points) Task 0a: Introduction 

Introduce the question to your audience. Suppose that they _could_ be interested in the question, but they don't have a deep background in the area. What is the question that you are addressing, why is it worth addressing, and what are you going to find at the completion of your analysis. Here are a few resource that you might use to start this motivation. 

- [Wikipedia](https://en.wikipedia.org/wiki/Keeling_Curve)
- [First Publication](./background/keeling_tellus_1960.pdf)
- [Autobiography of Keeling](./background/keeling_annual_review.pdf)

The question we are addressing is the investigation of trends in global carbon dioxide (CO2) concentrations. The CO2 concentration data comes from the Mauna Loa Observatory, which has been collecting these measurements since 1958. This data has been instrumental in the study of climate change due to the greenhouse effect. Our study seeks to quantify these trends, both in terms of levels and growth rates, and to forecast future CO2 concentrations based on these trends. By the completion of our analysis, we aim to provide a comprehensive understanding of the CO2 data and its characteristics, fit suitable statistical models to this data, and make informed predictions about future atmospheric CO2 levels.


## (3 points) Task 1a: CO2 data
Conduct a comprehensive Exploratory Data Analysis on the `co2` series. This should include (without being limited to) a [description of how, where and why] (https://gml.noaa.gov/ccgg/about/co2_measurements.html) the data is generated, a thorough investigation of the trend, seasonal and irregular elements. Trends both in levels and growth rates should be discussed (consider expressing longer-run growth rates as annualized averages).

What you report in the deliverable should not be your own process of discovery, but rather a guided discussion that you have constructed so that your audience can come to an understanding as succinctly and successfully as possible. This means that figures should be thoughtfully constructed and what you learn from them should be discussed in text; to the extent that there is _any_ raw output from your analysis, you should intend for people to read and interpret it, and you should write your own interpretation as well.

```{r}
plot(co2, main="Atmospheric CO2 Concentration", xlab="Year", ylab="CO2 (ppm)")

co2_decomp <- decompose(co2)
plot(co2_decomp)

acf(co2)
pacf(co2)

```


## (3 points) Task 2a: Linear time trend model

Fit a linear time trend model to the `co2` series, and examine the characteristics of the residuals. Compare this to a quadratic time trend model. Discuss whether a logarithmic transformation of the data would be appropriate. Fit a polynomial time trend model that incorporates seasonal dummy variables, and use this model to generate forecasts to the year 2020. 

```{r}
linear_model <- lm(co2 ~ time(co2))
summary(linear_model)
plot(residuals(linear_model))
```

```{r}
quadratic_model <- lm(co2 ~ time(co2) + I(time(co2)^2))
summary(quadratic_model)
plot(residuals(quadratic_model))
```
```{r}
# Time variable
co2_df <- data.frame(co2 = as.numeric(co2), time = as.numeric(time(co2)))

# Seasonal dummy variables
co2_df$month <- cycle(co2)
```


```{r}
seasonal_model <- lm(co2 ~ time + I(time^2) + factor(month), data = co2_df)
```


```{r}
future_time <- seq(max(time(co2)), 2020, by = 1/12)
future_month <- cycle(ts(future_time, start = 1959, freq = 12))
future_df <- data.frame(time = future_time, month = future_month)
pred <- predict(seasonal_model, newdata = future_df)
```


```{r}
# Plot the data and forecasts
plot(co2, main = "CO2 Concentration with Forecasts to 2020")
lines(future_time, pred, col = "blue")
```

## (3 points) Task 3a: ARIMA times series model 

Following all appropriate steps, choose an ARIMA model to fit to the series. Discuss the characteristics of your model and how you selected between alternative ARIMA specifications. Use your model (or models) to generate forecasts to the year 2022. 

```{r}
arima_model <- auto.arima(co2)
summary(arima_model)
forecast(arima_model, h=24*12) 
```


## (3 points) Task 4a: Forecast atmospheric CO2 growth 

Generate predictions for when atmospheric CO2 is expected to be at [420 ppm](https://research.noaa.gov/article/ArtMID/587/ArticleID/2764/Coronavirus-response-barely-slows-rising-carbon-dioxide) and 500 ppm levels for the first and final times (consider prediction intervals as well as point estimates in your answer). Generate a prediction for atmospheric CO2 levels in the year 2100. How confident are you that these will be accurate predictions?

```{r}
# Return the date that the given series first hit x num
first.date <- function(series, num) {
  return(dates[which(series > num)[1]])
}

forecast.1997 <- forecast(arima_model, h = 12 * 117, level = c(95))
dates <- rownames(data.frame(forecast.1997))

# When CO2 is expected to hit 420ppm
prediction.420.lower <- first.date(forecast.1997$lower, 420)
prediction.420.expected <- first.date(forecast.1997$mean, 420)
prediction.420.upper <- first.date(forecast.1997$upper, 420)

# When CO2 is expected to hit 500ppm
prediction.500.lower <- first.date(forecast.1997$lower, 500)
prediction.500.expected <- first.date(forecast.1997$mean, 500)
prediction.500.upper <- first.date(forecast.1997$upper, 500)

# Expected CO2 levels in the year 2100
prediction.2100.index <- match("Jan 2100", dates)
prediction.2100.lower <- forecast.1997$lower[prediction.2100.index]
prediction.2100.expected <- forecast.1997$mean[prediction.2100.index]
prediction.2100.upper <- forecast.1997$upper[prediction.2100.index]
```


```{r}
forecast_df <- data.frame(
  date = as.Date(seq(from = as.Date("1998-01-01"), by = "month", length.out = 12 * 117)),
  CO2 = forecast.1997$mean,
  lower = forecast.1997$lower,
  upper = forecast.1997$upper
)
```


```{r}
# Generate plot
ggplot(forecast_df, aes(x = date)) +
  geom_ribbon(aes(ymin = X95., ymax = X95..1), fill = "grey80", alpha = 0.5) +
  geom_line(aes(y = CO2), color = "darkblue") +
  geom_hline(yintercept = 420, linetype = "dashed", color = "red") +
  geom_hline(yintercept = 500, linetype = "dashed", color = "orange") +
  labs(
    x = "Year",
    y = "CO2 Concentration (ppm)",
    title = "Forecasted CO2 Concentration",
    subtitle = "Red line: 420 ppm, Orange line: 500 ppm"
  ) +
  theme_minimal()
```

# Report from the Point of View of the Present 

One of the very interesting features of Keeling and colleagues' research is that they were able to evaluate, and re-evaluate the data as new series of measurements were released. This permitted the evaluation of previous models' performance and a much more difficult question: If their models' predictions were "off" was this the result of a failure of the model, or a change in the system? 

## (1 point) Task 0b: Introduction 

In this introduction, you can assume that your reader will have **just** read your 1997 report. In this introduction, **very** briefly pose the question that you are evaluating, and describe what (if anything) has changed in the data generating process between 1997 and the present. 

## (3 points) Task 1b: Create a modern data pipeline for Mona Loa CO2 data.

The most current data is provided by the United States' National Oceanic and Atmospheric Administration, on a data page [[here](https://gml.noaa.gov/ccgg/trends/data.html)]. Gather the most recent weekly data from this page. (A group that is interested in even more data management might choose to work with the [hourly data](https://gml.noaa.gov/aftp/data/trace_gases/co2/in-situ/surface/mlo/co2_mlo_surface-insitu_1_ccgg_HourlyData.txt).) 

Create a data pipeline that starts by reading from the appropriate URL, and ends by saving an object called `co2_present` that is a suitable time series object. 

```{r pull data from the global monitoring laboratory}
co2_url <- "https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_weekly_mlo.csv"
co2_month_url <- "https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_mm_mlo.csv"

co2_present <- read.csv(co2_url, comment.char = "#") %>%
  mutate(index = yearweek(lubridate::make_datetime(year, month, day))) %>%
  mutate(average = ifelse(average == -999.99, NA, average)) %>%
  fill(average) %>%
  tsibble(index = index)
co2_present_month <- read.csv(co2_month_url, comment.char = "#") %>%
  mutate(index = yearmonth(lubridate::make_datetime(year, month))) %>%
  tsibble(index = index)
```


Conduct the same EDA on this data. Describe how the Keeling Curve evolved from 1997 to the present, noting where the series seems to be following similar trends to the series that you "evaluated in 1997" and where the series seems to be following different trends. This EDA can use the same, or very similar tools and views as you provided in your 1997 report. 

```{r}
co2_present %>%
  ggplot(aes(x = index, y = average)) +
  geom_line() +
  ggtitle("Weekly CO2 levels from start to present") +
  xlab("Year") +
  ylab("CO2 level")

# Plot the monthly series
co2_present_month %>%
  ggplot(aes(x = index, y = average)) +
  geom_line() +
  ggtitle("Monthly CO2 levels from start to present") +
  xlab("Year") +
  ylab("CO2 level")


# Decompose the time series to see trend, seasonality and residuals. The 'stl' function is used because 'decompose' won't work on 'tsibble' objects.
# Decompose the weekly series
co2_present_decomposed <- co2_present %>%
  model(STL(average ~ season(window = "periodic"))) %>%
  components()

autoplot(co2_present_decomposed)

# Decompose the monthly series
co2_present_month_decomposed <- co2_present_month %>%
  model(STL(average ~ season(window = "periodic"))) %>%
  components()

autoplot(co2_present_month_decomposed)
co2_present_acf <- ACF(co2_present, average)
autoplot(co2_present_acf)

co2_present_pacf <- PACF(co2_present, average)
autoplot(co2_present_pacf)

co2_present_month_acf <- ACF(co2_present_month, average)
autoplot(co2_present_month_acf)

co2_present_month_pacf <- PACF(co2_present_month, average)
autoplot(co2_present_month_pacf)


```

## (1 point) Task 2b: Compare linear model forecasts against realized CO2

Descriptively compare realized atmospheric CO2 levels to those predicted by your forecast from a linear time model in 1997 (i.e. "Task 2a"). (You do not need to run any formal tests for this task.) 

```{r}
data.1997 <- data.frame(index = 1:nrow(co2_ts), month = factor(month(co2_ts$index)), value = co2_ts$value)
```

```{r}
# Generate newdata to predict on
data.1997.max_month <- max(data.1997$index)
data.future.max_month <- data.1997.max_month + (12 * (2023 - 1997))
data.future <- data.frame(index = (data.1997.max_month + 1):data.future.max_month)

# Create forecasts
model.linear.forecast <- predict(model.linear, data.future, interval = "prediction", level = 0.99)
model.linear.predicted <- cbind(data.future, model.linear.forecast) %>%
  mutate(value = fit, index = yearmonth(seq(as.Date("1998/1/1"), as.Date("2023/12/1"), by = "month"))) %>%
  select(value, index) %>%
  head(-6) # remove the last 6 months

ggplot() +
  geom_line(aes(x = index, y = average), color = 'black', data = co2_present_month) +
  geom_line(aes(x = index, y = value), color = 'steelblue', data = model.linear.predicted) +
  labs(
    title = TeX(r'(Monthly Mean $CO_2$ with Forecasted Values)'),
    subtitle = 'Forecasted levels in blue',
    x = 'Month',
    y = TeX(r'($CO_2$ parts per million)')
  )
```

## (1 point) Task 3b: Compare ARIMA models forecasts against realized CO2  

Descriptively compare realized atmospheric CO2 levels to those predicted by your forecast from the ARIMA model that you fitted in 1997 (i.e. "Task 3a"). Describe how the Keeling Curve evolved from 1997 to the present. 

```{r}
model.arima.predicted <- data.frame(forecast(model.arima, h = (12 * 25) + 6, level = c(0)))
model.arima.predicted <- data.frame(
  index = yearmonth(rownames(model.arima.predicted)),
  average = model.arima.predicted$Point.Forecast
)

ggplot() +
  geom_line(aes(x = index, y = average), color = 'steelblue', data = model.arima.predicted) +
  geom_line(aes(x = index, y = average), color = 'black', data = co2_present_month) +
  labs(
    title = TeX(r'(Monthly Mean $CO_2$ with Forecasted Values)'),
    subtitle = 'Forecasted levels in blue',
    x = 'Month',
    y = TeX(r'($CO_2$ parts per million)')
  )
```


## (3 points) Task 4b: Evaluate the performance of 1997 linear and ARIMA models 

In 1997 you made predictions about the first time that CO2 would cross 420 ppm. How close were your models to the truth? 

After reflecting on your performance on this threshold-prediction task, continue to use the weekly data to generate a month-average series from 1997 to the present, and compare the overall forecasting performance of your models from Parts 2a and 3b over the entire period. (You should conduct formal tests for this task.) 

```{r}
co2_monthly_avg <- co2_present %>%
  mutate(month_year = paste(year, month, sep = "-")) %>%
  group_by(month_year) %>%
  summarise(avg_co2 = mean(average, na.rm = TRUE))

co2_monthly_avg$month_year <- as.Date(paste(co2_monthly_avg$month_year, "01", sep = "-"))
co2_monthly_avg_ts <- ts(co2_monthly_avg$avg_co2, start = c(1997, 1), frequency = 12)

#Comparison of forecasted and actual values
comparison <- data.frame(
  date = time(co2_monthly_avg_ts),
  actual = as.numeric(co2_monthly_avg_ts),
  predicted = c(forecast.1997$mean, rep(NA, length(co2_monthly_avg_ts) - length(forecast.1997$mean)))
)

head(comparison)
```


```{r}
#Error metrics
comparison <- comparison %>%
  mutate(error = actual - predicted,
         MAE = abs(error),
         RMSE = error^2)

# Calculate overall MAE and RMSE
MAE <- mean(comparison$MAE, na.rm = TRUE)
RMSE <- sqrt(mean(comparison$RMSE, na.rm = TRUE))

# 3. Formal test: Mincer-Zarnowitz regression
comparison <- na.omit(comparison)
MZ_reg <- lm(actual ~ predicted + 0, data = comparison)

MAE
RMSE
MZ_reg
```
In this case, an MAE of 104.9443 means that our model, on average, makes an error of about 104.9443 units in its predictions. An RMSE of 112.3219 means that our model's predictions are, on average, approximately 112.3219 units away from the actual values.

The linear regression is modeling the actual CO2 level as a function of the predicted CO2 level. The coefficient for the predicted variable is 0.762291. This means that for each unit increase in the predicted CO2 level, we would expect an average increase of about 0.76 units in the actual CO2 level. The p-value is extremely small (< 2.2e-16), meaning this effect is statistically significant. The R-squared is 0.9936, indicating that 99.36% of the variance in the actual CO2 levels is explained by the predicted levels, which is quite high.

However, there are significant residuals, ranging from -47.105 to 55.018, with a median of 3.428. This suggests that while the predicted values explain much of the variance, they systematically under- or overestimate the actual CO2 levels by up to about 55 units.

```{r}
#Forecast performance: Diebold-Mariano test
DM_test <- dm.test(comparison$actual, comparison$predicted, alternative = "two.sided")
DM_test
```
The DM test is a procedure for comparing forecast accuracy. The null hypothesis is that the two methods (in this case, the actual and predicted values) have the same forecast accuracy.

The DM statistic is -83.65 and the p-value is less than 2.2e-16. This extremely low p-value leads us to reject the null hypothesis, implying that there is a statistically significant difference between the forecast accuracy of the actual and predicted CO2 levels. Given that the DM statistic is negative, it indicates that the actual values have a higher forecast accuracy than the predicted values.

https://eranraviv.com/volatility-forecast-evaluation-in-r/


```{r}
ggplot(data = comparison, aes(x = predicted, y = actual)) +
  geom_point(color = 'blue') +
  geom_smooth(method = 'lm', color = 'red') +
  labs(title = "Actual vs. Predicted CO2 Levels", x = "Predicted", y = "Actual") +
  theme_minimal()

```
In summary, while the model does a good job of explaining the variance in actual CO2 levels (as indicated by the high R-squared value), there are significant differences between the actual and predicted values, suggesting that the model may be systematically under- or overestimating the actual levels. Additionally, the DM test suggests that the actual values have a higher forecast accuracy than the predicted values.

## (4 points) Task 5b: Train best models on present data

Seasonally adjust the weekly NOAA data, and split both seasonally-adjusted (SA) and non-seasonally-adjusted (NSA) series into training and test sets, using the last two years of observations as the test sets. For both SA and NSA series, fit ARIMA models using all appropriate steps. Measure and discuss how your models perform in-sample and (psuedo-) out-of-sample, comparing candidate models and explaining your choice. In addition, fit a polynomial time-trend model to the seasonally-adjusted series and compare its performance to that of your ARIMA model.

```{r}
co2_ts <- ts(co2_present$average, frequency = 52)

# Determine the seasonal component 
co2_stl <- stl(co2_ts, s.window = "periodic")

# Seasonal adjustment
co2_present_sa <- data.frame(
  Date = co2_present$Date,
  average_sa = co2_stl$time.series[, "seasonal"]
)
```


```{r}
train_sa <- co2_present_sa %>% filter(Date < as.Date("2021-01-01"))
test_sa <- co2_present_sa %>% filter(Date >= as.Date("2021-01-01"))

train_nsa <- co2_present %>% filter(Date < as.Date("2021-01-01"))
test_nsa <- co2_present %>% filter(Date >= as.Date("2021-01-01"))
```


```{r}
model_sa <- auto.arima(train_sa$average_sa)
model_nsa <- auto.arima(train_nsa$average)
```

```{r}
summary(model_sa)
```


```{r}
summary(model_nsa)
```
```{r}
pred_sa <- forecast(model_sa, h = length(test_sa$average_sa))
pred_nsa <- forecast(model_nsa, h = length(test_nsa$average))
```

## TO:DO - Compare Models (SA & NSA) and analyze error metrics

```{r}
train_sa$Date <- as.numeric(train_sa$Date)
test_sa$Date <- as.numeric(test_sa$Date)

model_poly <- lm(average_sa ~ poly(Date, 4), data = train_sa)
poly_pred <- predict(model_poly, newdata = test_sa)
summary(model_poly)
```

## (3 points) Task Part 6b: How bad could it get?

With the non-seasonally adjusted data series, generate predictions for when atmospheric CO2 is expected to be at 420 ppm and 500 ppm levels for the first and final times (consider prediction intervals as well as point estimates in your answer). Generate a prediction for atmospheric CO2 levels in the year 2122. How confident are you that these will be accurate predictions?
