---
title: 'Statistical Methods for Discrete Response, Time Series, and Panel Data (W271): Lab 2'
geometry: margin=1in
output:
  github_document: default
---

```{r load packages, echo = FALSE, message = FALSE}
library(tidyverse)
library(tsibble)
library(latex2exp)

# Already presented in the course thus far
library(feasts)
library(forecast)
library(tseries)
library(lubridate)

theme_set(theme_minimal())
knitr::opts_chunk$set(dpi=1000)
```

# C02 Emissions

## Part 1

### Introduction

$CO_2$ levels have been recorded at the Mauna Loa observatory for over 40 years. At this time the data seem to show an alarming trend of increasing $CO_2$ levels year over year. This is alarming because $CO_2$ contributes to the "greenhouse effect", where certain gasses collect in the Earth's atmosphere and trap heat from leaving the Earth. As $CO_2$ levels increase we expect the Earth's temperature to increase with it. While the exact effects of the change this will have on the Earth's environment remain to be seen, expected changes include but are not limited to the following:
- Heat waves
- Drought
- Rising sea levels
With the data at hand, it is imperative that we discover whether we have enough evidence to show that this recent rise in $CO_2$ levels is the result of a larger trend or could be explained by natural variation. If this trend is confirmed then it could pave the way to future research on ways to measure and address the adverse effects and causes of this rise in $CO_2$. This report will look into the existence of this larger trend of rising $CO_2$ levels and, if it exists, will also report on the magnitude of the rise as well as project future $CO_2$ levels.

```{r plot the keeling curve, echo = FALSE}
tsibble::as_tsibble(co2) %>%
  ggplot() + 
  aes(x=index, y=value) + 
  geom_line(color = 'steelblue') +
  labs(
    title = TeX(r'(Monthly Mean $CO_2$)'),
    subtitle = 'The "Keeling Curve"',
    x = 'Month and Year',
    y = TeX(r'($CO_2$ parts per million)')
  )
```

### CO2 Data

#### Mauna Loa Site

As stated above, the data we will be using for this analysis is the $CO_2$ measurements from a laboratory at Mauna Loa, Hawaii. While there are other laboratories that collect $CO_2$ measurements, the Mauna Loa site has been collected $CO_2$ longer than any other site in the world which will give us the most data to work with as we conduct this analysis. The Mauna Loa site is also unique in that it is representative of air for the entire Northern Hemisphere due to its altitude and is not usually affected by nearby vegetation as the site is surrounded be lava flows.

The Mauna Loa data is frequently used because of the amount *and* quality of the data collected. Specifically, this dataset contains accurate and robust measurements of the number of $CO_2$ molecules per million in a cubic meter of *dry* air. The term **concentration** may be used for familiarity but it should be stated that this is not the preferred term as the concentration of $CO_2$ may be affected be a number of factors unrelated to how much $CO_2$ is actually in the world's atmosphere at a given moment.

This site measures the concentration of $CO_2$ by funnelling air through a cold chamber (to eliminate the effect of humidity) and then measuring how much infrared radiation is absorbed by the $CO_2$ in the chamber. Because $CO_2$ naturally absorbs infrared radiation, a higher density of $CO_2$ molecules will absorb more radiation. The researchers at the Mauna Loa site take great care to continually calibrate their equipment multiple times a day. In addition, the researchers are careful to account for any outside factors that may effect measurements such as the diurnal wind flow patterns present on Mauna Loa. Altogether, we can be confident that the data recorded at Mauna Loa is representative of global $CO_2$ concentrations.

#### Data Introduction

TODO: Fill this out with the data EDA

### Polynomial Time Trend Models

```{r}
ts = as_tsibble(co2)
ts
df = data.frame(1:nrow(co2_ts), month = factor(month(ts$index)), CO2_ppm = ts$value)
df
```
###2A: Linear Model
```{r}
linear_model = lm(CO2_ppm ~ month, data = df)
checkresiduals(linear_model)
summary(linear_model)
```

The linear model is a poor model because the p-value of the months is > 0.05. Therefore we lack evidence to reject the null hypothesis and are not able to conclude that there is a linear relationship between Looking at the rsidual plots, we can see a trend in the residuals. The ACF plot also shows strong autocorrelation, violating the assumption of a linear model. Therefore, a linear model does not sufficiently describe the relationship.

### 2b: ARIMA Model

```{r}

arima_model = co2 %>% auto.arima(ic="aicc", seasonal = FALSE)
arima_model

checkresiduals(arima_model)
summary(arima_model)
```

(
This ARIMA model captures the pattern pretty well except for the seasonality (non-seasonal model). The residuals look like a white noise process, which is desirable.

forecasting:
```{r}
predicted = forecast(arima_model,600,c(95))
predicted %>%
  autoplot() +
  labs(
    title = "CO2 and Forecast",
    x = "Month",
    y = "CO2_PPM"
  )
```

```{r}
forecast_1997 = forecast(arima_model, h = 2000, level = c(95))

expected_400 = rownames(data.frame(forecast_1997))[which(forecast_1997$mean> 400)[1]]
expected_500 = rownames(data.frame(forecast_1997))[which(forecast_1997$mean> 500)[1]]

p_2030 = forecast_1997$mean[match("Jan 2030", rownames(data.frame(forecast_1997)))]
p_2050 = forecast_1997$mean[match("Jan 2050", rownames(data.frame(forecast_1997)))]
p_2100 = forecast_1997$mean[match("Jan 2100", rownames(data.frame(forecast_1997)))]

expected_400
p_2100
```



## Part 2



### 1b

```{r}
co2_url <- "https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_mm_mlo.csv"
co2_present <- read.csv(co2_url, comment.char = "#") %>%
  mutate(index = yearmonth(lubridate::make_datetime(year, month))) %>%
  tsibble(index = index)
```

### 2b Model Comparisons



#### 3b ARIMA

```{r}

```

### 4b evaluations
